Multiple Linear Regression on Sleep Data

### 2025-05-06 paper/paper.qmd
paper work

sleep$stress_level <- as.factor(sleep$stress_level)
sleep$quality_of_sleep <- as.factor(sleep$quality_of_sleep)

@@

knitr::kable(model_stats, caption = "Full Linear Model Outcome")
```

@@

As far as assumptions go heteroskadcity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality.

```{r}
kable(vif(full_model))
```

@@

>>>>

sleep$stress_level <- as.factor(sleep$stress_level)
sleep$gender <- as.factor(sleep$gender)
sleep$occupation <- as.factor(sleep$occupation)
sleep$bmi_category <- as.factor(sleep$bmi_category)
sleep$sleep_disorder <- as.factor(sleep$sleep_disorder)

@@

knitr::kable(model_stats, caption = "Full Linear Model Outcome")
```
We also saw all coefficients have standard error less than one.

@@

As far as assumptions go heteroskadcity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality. We will explore multicollinearity and outliers next.

```{r}
kable(vif(full_model))
```
`age` and `quality_of_sleep` have $\text{GVIF}^{2\cdot\text{DF}} > 5$. We may want to explore Lasso to deal with this.

@@

```{r}
#Values of Interest
leverage <- hatvalues(full_model)
studentized <- rstudent(full_model)
dffits_vals <- dffits(full_model)
cooks_d <- cooks.distance(full_model)

diagnostics <- data.frame(Leverage = leverage,
                          Studentized = studentized,
                          DFFITS = dffits_vals,
                          Cooks_Distance = cooks_d)

#Thresh hold
n <- nrow(sleep)
p <- length(coef(full_model)) - 1
leverage_threshold <- 2 * (p + 1) / n
studentized_threshold <- 2
cooks_threshold <- 1

# Find points exceeding thresholds
high_leverage <- which(leverage > leverage_threshold)
high_studentized <- which(abs(studentized) > studentized_threshold)
high_cooks <- which(cooks_d > cooks_threshold)

# Unique indices that exceed any threshold
suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))

sleep[suspicious_indices,]
```

```{r}
leverage <- hatvalues(full_model)
studentized <- rstudent(full_model)
dffits_vals <- dffits(full_model)
cooks_d <- cooks.distance(full_model)

diagnostics <- data.frame(Leverage = leverage,
                          Studentized = studentized,
                          DFFITS = dffits_vals,
                          Cooks_Distance = cooks_d)

n <- nrow(sleep)
p <- length(coef(full_model)) - 1
leverage_threshold <- 2 * (p + 1) / n
studentized_threshold <- 2
cooks_threshold <- 1

# Find points exceeding thresholds
high_leverage <- which(leverage > leverage_threshold)
high_studentized <- which(abs(studentized) > studentized_threshold)
high_cooks <- which(cooks_d > cooks_threshold)

# Unique indices that exceed any threshold
suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))

# Return threshold values, diagnostic preview, and rows flagged as potential outliers
list(
  leverage_threshold = leverage_threshold,
  studentized_threshold = studentized_threshold,
  cooks_threshold = cooks_threshold,
  diagnostics_preview = head(diagnostics, 10),
  flagged_rows = diagnostics[suspicious_indices, ])

# Remove the suspicious rows
sleep_clean <- sleep[-suspicious_indices, ]

# Refit the model excluding suspicious rows 
full_model_clean <- lm(sleep_duration ~ ., data = sleep_clean)

# Plot diagnostics again on the cleaned model
par(mfrow = c(2, 2))
plot(full_model_clean)

# Comparison of model summaries
summary(full_model)        # Old model (with potential outliers)  R^2  = 0.9592
summary(full_model_clean)  # New model (without potential outliers)  R^2  = 0.9929

```
